### 2014-05-05 IEEE-Latex/Paper.tex
added first Paper draft

\documentclass[9pt,technote]{IEEEtran}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{algorithmic}
\usepackage{placeins}
\usepackage{hyperref}

\begin{document}

\begin{titlepage}
\begin{center}
\vfill{{\Large Active Queue Management}}
\end{center}

\end{titlepage}

\tableofcontents

\newpage

\section{Motivation}

\section{Congestion Control in Network Traffic}

Recommendations on Queue Management and Congestion Avoidance in the Internet					\url{http://www.hjp.at/doc/rfc/rfc2309.html}

Congestion Control in IP/TCP Internetworks														\url{http://www.hjp.at/doc/rfc/rfc896.html}

Congestion control in computer networks: issues and trends										\url{http://citeseerx.ist.psu.edu/viewdoc/summary?doi=10.1.1.122.4570}

Congestion control and traffic management in ATM networks: Recent advances and a survey			\url{http://www.sciencedirect.com/science/article/pii/0169755296000128}

TCP behavior with many flows																	\url{http://dl.acm.org/citation.cfm?id=852431&CFID=330551302&CFTOKEN=30669788}

Design considerations for supporting TCP with per-flow queueing									\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=659666&url=http\%3A%2F%2Fieeexplore.ieee.org\%2Fxpls%2Fabs_all.jsp\%3Farnumber%3D659666}

TCP and explicit congestion notification														\url{http://dl.acm.org/citation.cfm?id=205512&CFID=330551302&CFTOKEN=30669788}

\section{Definition and usage of Active Queue Managment}

Efficient Active Queue Management for Internet Routers											\url{http://www.researchgate.net/publication/2621818_Efficient_Active_Queue_Management_for_Internet_Routers}

Router Mechanisms to Support End-to-End Congestion Control										\url{http://citeseerx.ist.psu.edu/viewdoc/summary?doi=10.1.1.39.7772}

Sizing router buffers																			\url{http://dl.acm.org/citation.cfm?id=1015499}

Stochastic Modeling and the Theory of Queues													\url{http://www.gbv.de/dms/ilmenau/toc/018830102.PDF}

Analysis and simulation of a fair queueing algorithm											\url{http://dl.acm.org/citation.cfm?id=75248}

\section{Algorithms for Active Queue Management}

\subsection{Passive Techniques}
\subsubsection{CHOKe}
CHOKe - a stateless active queue management scheme for approximating fair bandwidth allocation	\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=832269&url=http\%3A%2F%2Fieeexplore.ieee.or0\%2Fxpls%2Fabs_all.jsp\%3Farnumber%3D832269}
\subsubsection{ECN}
ECN-Explicit Congestion Notification															http://www.hjp.at/doc/rfc/rfc3168.html
\subsection{Random Early Detection}
RED														Random early detection gateways for congestion avoidance										\url{http://ieeexplore.ieee.org/xpl/articleDetails.jsp?arnumber=251892&navigation=1}
	RED + Vorschläge für Architektur etc.					A study of active queue management for congestion control 										\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=832541&url=http\%3A\%2F\%2Fieeexplore.ieee.org\%2Fxpls\%2Fabs_all.jsp\%3Farnumber\%3D832541}

RED + Vorschläge für Architektur etc.					A study of active queue management for congestion control 										\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=832541&url=http\%3A%2F%2Fieeexplore.ieee.org\%2Fxpls%2Fabs_all.jsp\%3Farnumber%3D832541}

\subsection{Alternatives to RED}

\subsubsection{PI Controller}

PI														On designing improved controllers for AQM routers supporting TCP flows 							\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=916670&url=http\%3A%2F%2Fieeexplore.ieee.org\%2Fxpls%2Fabs_all.jsp\%3Farnumber%3D916670}

Vergleich RED, ARED, PI									The effects of active queue management on web performance										\url{http://dl.acm.org/citation.cfm?id=863986}

\subsubsection{Adaptive Virutal Queue}
Adaptive virtual Queue + (Vergleich zu RED, REM, PI)	Analysis and design of an adaptive virtual queue (AVQ) algorithm for active queue management	\url{http://dl.acm.org/citation.cfm?id=383069}

\subsubsection{BLUE}
BLUE													The BLUE active queue management algorithms														\url{http://dl.acm.org/citation.cfm?id=581869}

\section{Conclusion and Future Work}

\end{document}